\documentclass[12pt, a4paper, twoside, openright]{book}

\usepackage{vuwthesis} % sets up some local things, mostly the front page

%\usepackage{palatino} % sets palatino as the default font
\usepackage{fontspec}
\setmainfont{TeX Gyre Pagella}
\usepackage{url} % for typesetting urls

\usepackage[colorlinks=true, linkcolor=black, urlcolor=black,citecolor=black]{hyperref}

\usepackage{xspace}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{siunitx}
\usepackage[table]{xcolor}
\usepackage[bottom=3.0cm, left=3cm, right=3.0cm]{geometry}

\usepackage{pdflscape}
\usepackage{acronym}
%\renewcommand{\baselinestretch}{1.00}


\newcommand{\Ttwo} {\textit{T\textsubscript{2}}\xspace}
\newcommand{\SOtwo} {\textit{sO\textsubscript{2}}\xspace}
\newcommand{\TtwoO} {\textit{T\textsubscript{20}}\xspace}
\newcommand{\Kzero} {\textit{K\textsubscript{0}}\xspace}
\newcommand{\Gzero} {\textit{G\textsubscript{0}}\xspace}
\newcommand{\Otwo} {O\textsubscript{2}\xspace}
\newcommand{\Ntwo} {N\textsubscript{2}\xspace}
\newcommand{\COtwo} {CO\textsubscript{2}\xspace}
\newcommand{\TR} {\textit{T\textsubscript{R}}\xspace}
\newcommand{\Texc} {\textit{τ\textsubscript{ex}}\xspace}
\newcommand{\rc} {\textit{r\textsubscript{c}}\xspace}
\newcommand{\Tech} {\textit{t\textsubscript{ec}}\xspace}
\newcommand{\Rtwo} {\textit{R\textsubscript{2}}\xspace}
\newcommand{\Bzero} {\textit{B\textsubscript{0}}\xspace}
\newcommand{\Bone}  {\textit{B\textsubscript{1}}\xspace}
\newcommand{\Ttwostar} {\textit{T\textsubscript{2}\**}\xspace}
\newcommand{\Tone} {\textit{T\textsubscript{1}}\xspace}
\newcommand{\Hct} {\textit{Hct}\xspace}

\sisetup{separate-uncertainty = true}
%\sisetup{unit-color = blue}
%\sisetup{number-color = blue}

\begin{document}

\frontmatter
% Book style knows about front matter
% Report style doesn't so you need to set roman numbering etc yourself :-(

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Parameter Space Mapping for Blood Oxygenation Measurement with Low Field NMR}
\author{Dion Gary Thomas}

\subject{Physics}
\abstract{
Blood oxygenation is a critical physiological parameter for patient health.
The clinical importance of this parameter means that measurement of blood oxygenation is a routine part of care.
Magnetic resonance provides a way to measure blood oxygenation through the paramagnetic effect of deoxy-haemoglobin, which decreases the \Ttwo relaxation time of blood.
This effect has been well characterised at high fields (\SI{>1.5}{T}) for use in Magnetic Resonance Imaging, and it is a contributing factor to the Blood Oxygenation Level Dependent contrast used in functional MRI.
However there are relatively few studies of this effect at low magnetic fields, and these have only looked at extreme levels of oxygenation/deoxygenation.
To study this effect for potential application in a low-field device, we measured this effect to determine how factors such as oxygenation, field strength and CPMG echo time affect the \Ttwo of blood.

A continuous flow circuit, similar to a cardiopulmonary bypass circuit, was used to control parameters such as oxygen saturation and temperature, before the blood sample flowed into a variable field magnet (set at fields between 5-40 MHz), where a series of CPMG experiments with echo times ranging from \SIrange{1}{20}{ms} were performed to measure the \Ttwo.
Additionally, the oxygen saturation was continually monitored by an optical sensor, for comparison with the \Ttwo changes.
This allowed us to test the sensitivity of this effect at low fields.

These results show that at low fields, the \Ttwo relaxation change still follows the trends shown in the literature, with a dependence on \Bzero squared, and on the fraction of deoxyhaemoglobin squared.
Additionally, these results were also compared with two theoretical models for the dependence on echo time, which have previously been tested at high fields: the Luz-Meiboom equation, and the Jensen and Chandra model.
Both models gave good agreement with the data measured at low fields.
These experiments show that the \Ttwo changes in blood due to oxygenation are still visible at low field, and that this technique should be feasible in a low field device.


%Changes in blood oxygen saturation change the magnetic properties of red blood cells, generating an inhomogeneous magnetic field which decreases the \Ttwo relaxation time of the water protons in blood.
%This effect has been well characterised at imaging fields (\SI{>1.5}{T}), and has been applied to measure changes in blood oxygen saturation \textit{in-vivo}, but at the low fields required for portable NMR devices, there are only a handful of studies in the literature.
%These showed that the size of the \Ttwo change increases with the field strength squared, althuoguh these experiments at low fields only looked at extreme levels of oxygenation/deoxygenation.

%At these field strengths and echo times, both models provided good agreement with the experimental data, however the Jensen and Chandra model model was able to provide better fits to the data.
}
% Books don't normally have abstracts, and this is a bit of a hack

% Uncomment the appropriate degree
%\phd
\mscthesisonly
%\mscwithhonours
%\mscbothparts
% \otherdegree{DEGREE OR DIPLOMA NAME}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypersetup{pdftitle=\@title}



\maketitle

\include{chapters/acknowledge}
\addtocontents{toc}{\vskip-1.0cm}

\tableofcontents

\listoffigures
\listoftables
\include{chapters/acronyms}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% book style knows about mainmatter
% if you are using report style you will have to rest page numbering etc.
\mainmatter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% individual chapters included here

\include{chapters/introduction}
\include{chapters/background}
\include{chapters/exptsetup}
\include{chapters/stoppedflow}
%\include{chapters/continuousSO2setup}
\include{chapters/continuousSO2}
\include{chapters/diffmodels}
\include{chapters/conclusion}
\include{chapters/conference}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% and of course book style knows about backmatter
% \backmatter caused problems with appendices :-(
% and of course report style doesn't
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\bibliographystyle{ieeetr}
%\bibliographystyle{plain}
\bibliography{mscthesis}


\end{document}
