\chapter*{Acknowledgments}\label{ch:ack}
Firstly I need to thank my supervisors: Sergei Obruchkov, Shieak Tzeng and Petrik Galvosas for their excellent supervision, and for giving me the opportunity to pursue this research.
In particular I would like to thank them for their encouragement when the baby-MRI decided to choke a month before I was due to give a talk about results collected on it.
I would like to think that the panic and stress that caused was worth it.

Another big thanks goes to the other members and friends of the VUW NMR group over the last 2 years.
Thank you for making this such a fun place to work, and for all of your help and friendship.
To my other past supervisors, Marcel and Tim, thank you for introducing me to NMR and for letting me spend my honours year here.

Next, I would like to thank my fantastic friends and flatmates.
Thanks to them for putting up with my work stories, and for helping to keep me sane.
Of course, I must also thank Wilbur, who is somehow always excited to see me get home.

I would not have been able to complete this work without the care and support I have received from my family.
Thanks to my Mum and Dad for their encouragement, advice and support, and for treating me to weekend brunches.
And to my brother Adrian for his motivated assistance in acquiring research materials.
Thanks also to my family in the Philippines, especially my Lola, for their support and prayers.

I gratefully acknowledge funding support from the New Zealand Ministry of Business, Innovation and Employment, and the Australian and New Zealand Society for Magnetic Resonance for a travel grant to present this work at the ANZMAG conference.

I look forward to more experimenting in the future
